Cost- and Deadline-constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds

### 2012-01-19 paper/paper.tex
conclusions added

Summary:

 - new interesting problem was addressed and analyzed

 - dynamic and static strtegies were evaluated

 - new algorithms were developed

 - simulation studies were done

Lessons learned:

 - pros and cons of dynamic/static methods

 - problems faced and solved

 - observations about parallelization level, shape of the plan

Future work:

 - deal with uncertaineties of predictions

 - deal with heterogeneity and multiple clouds (ref to Gideon's and Jens'
 experiments on EC2 and FutureGrid)

 - add data transfers to cost models (ref to paper on data sharing)

>>>>

In this paper we addressed a new interesting and important problem of scheduling
and resource provisioning for scientific workflow ensembles on IaaS cloud
infrastructures. The problem is different from previous work on grid or utility
grid scheduling that cloud infrastructure gives more control over the resources,
so the resource provisioning plan can be adjusted to the application
requirements. Therefore the problem space becomes larger: it requires not
only to select the best scheduling of tasks to available resources, but also
select the best shape of resource provisioning plan. The formulation of a
problem as maximization of the number of prioritized workflows completed from
the ensemble requires also to take the decisions on admitting or rejecting workflows
based on their estimated resource demands. We believe that such formulation of
the bi-constrained problem is highly relevant since such
constraints are typically imposed on many real-world projects.

We have analyzed both static and dynamic scheduling approaches and
developed the SPDS, DPDS, WA-DPDS and SPSS algorithms realizing these
strategies. The algorithms were evaluated on ensembles of benchmark
workflows which represent typical real scientific applications. For the
purpose of evaluation we have developed a simulator that models the cloud
infrastructure and the tightly-coupled scheduler and provisioner modules of
workflow engine. The results of simulation studies indicate that the two algorithms (WA-DPDS and
SPSD) that take into account the information about the workflow structure and
estimation of task runtimes yield better results than the simple on-line
priority-based scheduling strategy with static (constant) resource provisioning
plan.

During our work, several interesting lessons were learned about the workflow
ensembles on cloud infrastructures. One result is that the simple strategy of
SPDS performs relatively well in terms of number of workflows completed and the
high resource utilization it can achieve. However, this method does not allow to
give any advance prediction about the result, so to see its performance the
whole ensemble needs to be simulated. On the other hand, the method may be
the only choice in the case no estimates about the task runtimes are available
in advance. It can be seen that the workflow-aware and static algorithms are
able to find better quality solutions in terms of the defined metrics: this the
result mainly of the admission algorithm implemented, which allow rejecting too
large workflows and thus saving resources for completing the less costly ones.

Another general observation is that although clouds provide such capabilities as
dynamic adding and removing of resources at runtime and high parallelization due
to ilusion of unlimited resources, it turns out that is recommended to avoid
using these capabilities when possible, in order to achieve high resource
utilization and cost effectiveness. This observation comes from the fact that
starting and stopping more VMs incurres a cost which can be amortized by running
each started VM as long as possible before shutdown. This effect was confirmed
in the case of SPSS algorithm for Montage and CyberShake workflows with small
budgets. Similarly, high degree of parallelization (large number of nodes) leads
also to less efficient resource utilization and in the case of clouds the drop
in parallel efficiency means also the increase of cost. This result, clearly
visible in Fig~\ref{fig:cost}, leads to multi-criteria decision problems and
related trade-offs. E.g. our simulation results can be used as hints to answer
the question how much time and money is required to actually complete all the
workflows from the given ansemble.

Another lesson learned from our studies is related to the shapes and granularity
of the tasks of workflows in the ensembles. The benchmarks workflows that we
tested have relatively large number of tasks (50--1000). Montage and CyberShake
workflows have fine-grained tasks of short execution times, in the order of
seconds and minutes. LIGO, Epigenomics and SIPHT have longer tasks: from minutes to hours.
Therefore the dynamic algorithms achieved better results on the fine grained
workflows, since the simple random-based allocation strategy leads to good
load-balancing and high resource utilization. On the other hand, when tasks are
longer, it becomes more beneficial to apply a static planning strategy of SPSS.

Our current study reveals multiple paths for future work. One interesting
question would be to answer how the static and dynamic algorithms perform in the
cases of increased uncertainities and failures, that have important influence
on overall system performance~\cite{Sakellariou2010,Dongarra2007}. Second
important goal would be to extend our application and infrastructure model to
include various data storage options available on clouds. The experimental study
on that subject~\cite{Juve2010} suggests that data demands of scientific
workflows have a high impact not only on execution time but also on costs on
commercial clouds. Finally, we would like to address the heterogeneity of the
infrastructure, as multiple instance types and providers, including private
and community clouds will make the problem more complex and
challenging~\cite{Marshall2010,vockler11,Juve2010}.
